---
title: "Capstone Project - Predicting Patient No-Shows Using Appointment Data"
output:
  pdf_document: default
  word_document: default
  html_document: default
---


```{r message = FALSE}
library(tidyverse)
library(lubridate)
library(caret)
library(randomForest)
library(GGally)

```

Read data and assign to _appointments_
```{r}
appointments <- read_csv("Final_Data.csv")
zipcodes <- read_csv("zipcodes.csv")
```




### Data Summary and Structure

```{r}
summary(appointments)
str(appointments)
head(appointments[,1:5])
head(appointments[,6:10])
head(appointments[,11:14])

# Check for NAs
sapply(appointments, function(x) sum(is.na(x)))
```
patient_distance variable has 972 NA values.  



###Data Dictionary

```{r}
variable_description <- c(
    "Dependent variable: kept or missed",
    "Appointment date",
    "Appointment time",
    "Appointment length in minutes",
    "Date appointment was scheduled",
    "Patient age",
    "Patient gender",
    "Billing type",
    "Number of prior missed appointments",
    "Number of prior kept appointments",
    "Patient distance from office in miles",
    "Office Zip Code - Anonymized",
    "Provider primary specialty code",
    "Reminder Call result")
variable <- colnames(appointments)
              
as_data_frame(cbind(c(1:14),variable, variable_description))
```


Will combine the appointment time and date into one variable, appt_datetime.
```{r}

appointments_2 <- appointments %>%
    mutate(appt_datetime = lubridate::mdy_hms(paste(appt_date, appt_time)))
appointments_2$date_scheduled <- as.POSIXct(appointments_2$date_scheduled, 
                                     format = "%m/%d/%y")
```

Calculating percent of missed appointments overall. Will first create a
logical variable _missed_, where 1 represents a missed appointment and 0
represents a kept appointment.
```{r}
appointments_2 <- appointments_2 %>%
    mutate(missed = ifelse(appointments_2$kept_status == "Missed", 1,0))
missed_rate <- mean(appointments_2$missed)
missed_rate
```

About 16% of the total appointments are missed.


###Data Exploration

#### patient_age

```{r}
ggplot(
  data = appointments_2, 
  mappng = aes(x = patient_age)
) +
  geom_histogram(
    mapping = aes(x = patient_age, col = kept_status, fill = kept_status),
    binwidth = 10)

ggplot(data = appointments_2) +
    geom_bar(
        mapping = aes(x = patient_age, fill = kept_status),
        position = "fill")
```

Ranges from 0-264, so there are obviously a few impossible values.
Ratio of missed appointments decreases with age in general.

Removing obervations of ages greater than 100, creating categorical age groups
and replotting.
```{r}
appointments_2 <- appointments_2 %>%
    filter(patient_age <= 100)

appointments_2 <- appointments_2 %>%
    mutate(age_cat = cut(patient_age, breaks = c(-1, 10, 20, 30, 40, 50, 60, 70, 101),
                         labels = c("0-10", "10-20", "20-30", "20-40", "40-50",
                                    "50-60", "60-70", "Over 70")))


ggplot(appointments_2, aes(x = age_cat, group = kept_status, col = kept_status,
                           fill = kept_status)) +
    stat_count()

ggplot(data = appointments_2) +
    geom_bar(mapping = aes(x = age_cat, fill = kept_status), position = "fill")

ggplot(data = appointments_2, aes(x = kept_status, y = patient_age)) +
    geom_boxplot()
```


#### billing_type
```{r}
table(appointments_2$billing_type)

ggplot(data = appointments_2) +
    geom_bar(mapping = aes(x = billing_type, fill = kept_status), position = "fill")

```

Only one row has _To Be Assigned_ value and will just be removed
There is a minor difference between billing types. DMAP has a higher proportion
of missed appointments

```{r}
appointments_2 <- subset(appointments_2,
                           appointments_2$billing_type != "To Be Assigned")
```


#### appt_datetime

Creating new _hour_ variable and plot by hour
```{r}
appointments_2 <- appointments_2 %>%
    mutate(hour = lubridate::hour(appointments_2$appt_datetime))
    
ggplot(data = appointments_2, 
       aes(x = hour, group = kept_status, col = kept_status, fill = kept_status)) +
    geom_histogram(binwidth = 1)

ggplot(data = appointments_2) +
    geom_bar(mapping = aes(x = hour, fill = kept_status), position = "fill")

```

Ranges from 00:00:00 to 21:00:00.
More appointments are missed in the early morning, late afternoon and early
evening, and around lunchtime, however, there are very few appointments at
these times. During main scheduling periods, the variation is less significant.


#### remind_call_result
```{r}
table(appointments_2$remind_call_result)
```
Low counts of "Answered - Cancelled", "Answered - Reschedule", "Busy",
and "No Answer"
```{r}
ggplot(data = appointments_2) +
    geom_bar(aes(x = remind_call_result, fill = kept_status), position = "fill") +
    theme(axis.text.x = element_text(size = 8, angle = 45,
                                     hjust = 1, vjust = 1))
```

~65% of appointments with "Answered - Cancelled" and ~35% with
"Answered-Reschedule" still kept their appointments, however, very few
 observations in these categories.
 


#### provider_specialty
```{r}
table(appointments_2$provider_specialty)
```
48 missing values.
E and F have few observations.

```{r}
ggplot(data = appointments_2) +
    geom_bar(aes(x = provider_specialty, fill = kept_status), position = "fill") +
    theme(axis.text.x = element_text(size = 7))
```

C, D, and E provider specialties have lower proportion of missed appointments,


#### appt_length
```{r}
ggplot(appointments_2, aes(x = appt_length, group = kept_status, col = kept_status)) +
    geom_histogram(binwidth = 10)
ggplot(data = appointments_2) +
    geom_bar(mapping = aes(x = appt_length, fill = kept_status), position = "fill")
ggplot(data = appointments_2, aes(x = kept_status, y = appt_length)) +
    geom_boxplot()

```

#### patient_distance
```{r}
ggplot(appointments_2, aes(x = patient_distance, group = kept_status, col = kept_status)) +
    geom_histogram(binwidth = 10)
ggplot(data = appointments_2) +
    geom_bar(mapping = aes(x = patient_distance, fill = kept_status), position = "fill")
ggplot(data = appointments_2, aes(x = kept_status, y = patient_distance)) +
    geom_boxplot()

```
patient_distance is very right-skewed, therefore NA values will be replaced with median rather than mean.
```{r}
appointments_2$patient_distance <- appointments_2$patient_distance %>%
    tidyr::replace_na(median(appointments_2$patient_distance, na.rm = TRUE))

```


Create new variables

percent_missed = percent of prior appointments missed.
New represents represents first time appointments
appt_lead_time is the difference between the day the appointment was scheduled
and the day of the appointment.

```{r}
appointments_3 <- appointments_2 %>%
    mutate(percent_missed = prior_missed / (prior_missed + prior_kept)) %>%
    mutate(new = ifelse(appointments_2$prior_missed == 0 & appointments_2$prior_kept == 0, 1, 0)) %>%
    mutate(appt_lead_time = appt_datetime - date_scheduled) %>% 
    mutate(weekday = strftime(appt_datetime, "%A"))
```


Add city_size and county_code from zipcode data.
```{r}
appointments_3 <- dplyr::left_join(appointments_3, zipcodes, by = "office_zip")
str(appointments_3)
```

#### percent_missed

Create random subset and plot
```{r}
appointments_sample_05 <- appointments_3 %>%
    sample_frac(size = 0.05, replace = FALSE)

ggplot(data = appointments_sample_05, aes(x = percent_missed, y = missed)) +
    geom_point() +
    stat_smooth(method = "loess") +
    xlab("Percent of Prior Appointmens Missed") +
    ylab("Probability of Missed Appointment")
ggplot(data = appointments_3, aes(x = kept_status, y = percent_missed)) +
    geom_boxplot()
    
```



#### new

```{r}
table(appointments_3$new)

ggplot(data = appointments_3) +
    geom_bar(mapping = aes(x = new, fill = kept_status), position = "fill")
```

New patients have a very high percentage of kept appointments.
22k of 342k appointments are first-time, or about 6.4%

#### appt_lead_time

```{r}
table(appointments_3$new)

ggplot(data = appointments_3) +
    geom_bar(mapping = aes(x = appt_lead_time, fill = kept_status), position = "fill")
```

#### city_size
```{r}
table(appointments_3$new)

ggplot(data = appointments_3) +
    geom_bar(mapping = aes(x = city_size, fill = kept_status), position = "fill")
```

#### county_code
```{r}
table(appointments_3$new)

ggplot(data = appointments_3) +
    geom_bar(mapping = aes(x = county_code, fill = kept_status), position = "fill")
```



### Modeling

Create Modeling Data
```{r}
model_data <- appointments_3 #%>%
model_data$new <- as.factor(model_data$new)
model_data$percent_missed <- as.integer(model_data$percent_missed * 100)
#Replace NAs with mean
model_data$percent_missed <- model_data$percent_missed %>%
    tidyr::replace_na(mean(model_data$percent_missed, na.rm = TRUE))
factor_columns <- c("kept_status", "patient_gender", "billing_type", 
"office_zip", "provider_specialty", "remind_call_result", "weekday",
"county_code")
model_data[factor_columns] <- lapply(model_data[factor_columns], factor)
#Check for NAs
sapply(model_data, function(x) sum(is.na(x)))

str(model_data)

```


Divide model_data into train, validate, and test sets
```{r}

train <- model_data[1:205660,]
validate <- model_data[205660:274200,]
test <- model_data[274201:nrow(model_data),]

table(train$kept_status)
train2 <- train[168738:205660,]
table(train2$keptstatus)
train_kept <- train2[train2$kept_status == "Kept",]
train_missed <- train[train$kept_status == "Missed",]

train_balanced <- rbind(train_kept, train_missed)
table(train_balanced$kept_status)

```




### Logistic Regression Model

```{r eval = FALSE}

model1 <- caret::train(kept_status ~ age + remindresult + specialty + billtype + hour + percent_missed + length + gender + distance + new + leadtime + weekday + county, data = train_balanced, method = "glm")
model1$finalModel
confusionMatrix(model1)
##p_glm <- predict(glm, train)
#caret::confusionMatrix(p_glm, train$kept_status)
```



### Random Forest Model

Using randomForest Package
```{r}
rf <- randomForest(kept_status ~ patient_age + remind_call_result + provider_specialty + billing_type + hour + percent_missed + appt_length + patient_gender + patient_distance + new + appt_lead_time + weekday + county_code + city_size, data = train_balanced, ntree = 250)
#Takes about 30 seconds to run
```
```{r}
print(rf)
plot(rf)
varImpPlot(rf)
```


Using caret Package
```{r}
# Look at number of cvs and repeats for faster run-time
control <- caret::trainControl(method = "repeatedcv", number = 10, repeats = 3)
seed <- 7
metric <- "Accuracy"
set.seed(seed)
mtry <- 3
tunegrid <- expand.grid(.mtry = mtry)

#Train on subset to see how long it will take.  Takes ~ 1.5 hours
```

```{r eval = FALSE}
rftrain <- caret::train(keptstatus ~ age + remindresult + specialty + billtype + hour + percent_missed + length + gender + distance + new + leadtime + weekday + county, data = train_balanced, method = "rf", metric = metric, tuneGrid = tunegrid, trControl = control)
```

```{r eval = FALSE}
caret::confusionMatrix(rftrain)
```



```{r}
control <- caret::trainControl(method = "repeatedcv", number = 10, repeats = 3, search = "random")
seed <- 7
metric <- "Accuracy"
set.seed(seed)
mtry <- 3
tunegrid <- expand.grid(.mtry = mtry)
```

```{r eval = FALSE}
### Below code takes a long time to run, need to consider ways to shorten it
rftrain3 <- caret::train(keptstatus ~ age + remindresult + specialty + 
                         billtype + hour + percent_missed + length + gender +
                         distance + new + leadtime + weekday + county,
                         data = train_balanced, method = "rf", metric = metric,
                         tuneLength = 15, trControl = control)
```

```{r eval = FALSE}
print(rftrain2)
plot(rf)
varImpPlot(rf)
varUsed(rf)
p_rf <- predict(rf, test)
caret::confusionMatrix(p_rf, test$keptstatus)
```









